---
title: "HW2"
author: "Christina Dai cd3347"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

### Problem 1

Didn't figure it out :(

Clean the data in pols-month.csv:

```{r}
pols_clean = 
  read_csv("./data/pols-month.csv") %>% 
  separate(mon, into = c('year', 'month', 'day'), sep = '-') %>% 
  mutate(month = recode(month, `01` = "jan", `02` = "feb", `03` = "mar", `04` = "apr", `05` = "may", `06` = "jun", `07` = "jul", `08` = "aug", `09` = "sep", `10` = "oct", `11` = "nov", `12` = "dec")) %>% 
  mutate(president = recode(prez_gop, `0` = "dem", `1` = "gop", `2` = "gop")) %>% 
  select(everything(), -day, -prez_gop, -prez_dem)
```

Do the same for snp.csv:

```{r}
snp_clean = 
  read_csv("./data/snp.csv") %>% 
  separate(date, into = c('month', 'day', 'year'), sep = '/') %>%
  mutate(month = as.numeric(month)) %>% 
  arrange(year, month) %>% 
  mutate(month = recode(month, `1` = "jan", `2` = "feb", `3` = "mar", `4` = "apr", `5` = "may", `6` = "jun", `7` = "jul", `8` = "aug", `9` = "sep", `10` = "oct", `11` = "nov", `12` = "dec")) %>% 
  relocate(year, month)
```

Clean/tidy unemployment.csv

```{r}
unemployment_clean = 
  read_csv("./data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment_rate"
  ) %>% 
  mutate(year = as.character(year))
```

Merging snp into pols:

```{r}
snp_pols_merge = 
  left_join(pols_clean, snp_clean)
```


### Problem 2

Read and clean the Mr. Trash Wheel sheet:

```{r}
mr_trash_tidy = 
  read_excel("./data/202309 Trash Wheel Collection Data.xlsx", sheet = 1,
             range = "A2:N586") %>% 
  janitor::clean_names() %>% 
  mutate(homes_powered = (weight_tons * 500) / 30) %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(wheel_name = "mr_trash") %>% 
  relocate(wheel_name)
```


Do the same for the Professor Trash Wheel and Gwynnda datasets:

```{r}
prof_trash_tidy = 
  read_excel("./data/202309 Trash Wheel Collection Data.xlsx", sheet = 2,
             range = "A2:M108") %>% 
  janitor::clean_names() %>% 
  mutate(homes_powered = (weight_tons * 500) / 30) %>% 
  mutate(wheel_name = "prof_trash") %>% 
  relocate(wheel_name)
```

```{r}
gwynnda_tidy = 
  read_excel("./data/202309 Trash Wheel Collection Data.xlsx", sheet = 4,
             range = "A2:L157") %>% 
  janitor::clean_names() %>% 
  mutate(homes_powered = (weight_tons * 500) / 30) %>% 
  mutate(wheel_name = "gwynnda") %>% 
  relocate(wheel_name)
  gwynnda_cigs = gwynnda_tidy[1,10] + gwynnda_tidy[2,10] + gwynnda_tidy[3,10] + gwynnda_tidy[4,10] + gwynnda_tidy[5,10]
```


Combine all three datasets:

```{r}
trash_wheels_tidy = 
  bind_rows(mr_trash_tidy, prof_trash_tidy, gwynnda_tidy)
```

In the combined dataset, which includes data from Mr. Trash Wheel, Professor Trash Wheel, and Gwynnda Trash wheel, there are `r nrow(trash_wheels_tidy)` rows. One key variable in this dataset is `weight_tons`, which describes the weight in tons of trash collected by each dumpster receiving trash from each trash wheel. Another key variable is `homes_powered`, which describes the amount of homes that are powered by the electricity generated by the incineration of the trash collected. Other key variables include trash types, such as `polystyrene` and `glass bottles`. The dataset also includes information like year, month, and date. For available data, the total amount of trash collected by Professor Trash Wheel is `r sum(prof_trash_tidy$weight_tons)`. The total number of cigarette butts collected by Gwynnda in July of 2021 is `r gwynnda_cigs`.


### Problem 3

Import and clean the MCI baseline dataset:

```{r}
mci_baseline_tidy = 
  read_csv("./data/MCI_baseline.csv", skip = 1) %>% 
  janitor::clean_names() %>% 
  mutate(sex = recode(sex, `1` = "male", `0` = "female")) %>% 
  mutate(apoe4 = recode(apoe4, `1` = "carrier", `0` = "non-carrier")) %>% 
  filter(age_at_onset > current_age | age_at_onset == '.')

develop_mci = filter(mci_baseline_tidy, age_at_onset != '.')
carriers = filter(mci_baseline_tidy, apoe4 == "carrier")
```

The first step in this process is to read in the data using `read_csv`, skipping the first row because it does not contain any data. Then, after cleaning the names using `janitor::clean_names()`, we use `mutate()` to recode the `sex` variable and `apoe4` variable to be more specific. Lastly, using `filter`, we only include observations where `age_at_onset` is either greater than `current_age` or the null value of "." to ensure that inclusion criteria is met. 483 participants were recruited for this study, with `r nrow(mci_baseline_tidy)` meeting inclusion criteria. `r nrow(develop_mci)` participants developed MCI during the study. The average baseline age is `r mean(mci_baseline_tidy$current_age)`, while `r (nrow(carriers)/479) * 100` is the proportion of women in the study that are APOE4 carriers.


Now, do the same for the MCI amyloid dataset:

```{r}
mci_amyloid_tidy = 
  read_csv("./data/mci_amyloid.csv", skip = 1) %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    baseline:time_8,
    names_to = "record_time",
    names_prefix = "time_",
    values_to = "amyloid_ratio"
  ) %>% 
  rename("id" = study_id)
```

The first step here is the same as before: read the data using `read_csv`. Then use `janitor::clean_names()` to standardize the variable names. After that, we need to tidy our recording times by using `pivot_longer`, which will combine all the recording times into one variable, `record_time`, and amyloid values into `amyloid`. Now, each observation corresponds to the amyloid value at each time of recording, including baseline.

Now, check to see if some participants appear only in either dataset:

```{r}
anti_join(mci_baseline_tidy, mci_amyloid_tidy, by = "id")
```

There are 8 participants who don't exist in both datasets: IDs 14, 49, 92, 179, 268, 304, 389, and 412. These participants only exist in the baseline dataset, and were likely lost to follow-up.

Now, combine datasets together with only particpants that exist in both datasets:

```{r}
mci_combined = 
  inner_join(mci_baseline_tidy, mci_amyloid_tidy, by = "id")
```

The resulting dataset after combining the baseline and amyloid datasets has both the demographic variables and amyloid values in one complete dataset. It has `r nrow(mci_combined)` rows and `r ncol(mci_combined)` columns. Each amyloid ratio measured at the various points throughout the study gets its own observation, with 5 total observations per participant.

Exporting the combined data to a csv file:

```{r}
write_csv(mci_combined, "./data/mci_combined.csv")
```




